---
title: "Basic Statistics in R"
author: "ECO 6416"
date: "`r Sys.Date()`"
output: 
    pdf_document:
      toc: true
      number_sections: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

Here are all the packages needed to get started.

```{r packages}
library(gt)
library(tidyverse)
library(gtsummary)
library(plotly)


sessionInfo()
```


# Univariate Analysis

In univariate analysis, we look at a single variable and describe 3 different things:

- Center
- Shape
- Spread

## Center

Helps explain where the middle of the data is. This can be measured in 3 main ways. 

### Mean

```{r}
grades <- c(78,79,80,81,82)

mean(grades)
```

### Median 

```{r}
median(grades)
```
### Mode

There isn't an easy way of doing this, so I created a function instead.

```{r}
getModes <- function(x) {
  ux <- unique(x)
  tab <- tabulate(match(x, ux))
  ux[tab == max(tab)]
}


getModes(grades)
```

Since there is no value that occurs most frequently, they all show. Mode is rarely used.

## Shape 



# Optional - Fancier Output

Check out this fun stuff! Makes things look much cleaner.

## Summary Statistics

Using our classic `mtcars` dataset.

```{r}
mtcars%>% select(mpg, cyl,hp) %>% 
  tbl_summary(statistic = list(all_continuous() ~ c("{mean} ({sd})",
                                                    "{median} ({p25}, {p75})",
                                                    "{min}, {max}"),
                              all_categorical() ~ "{n} / {N} ({p}%)"),
              type = all_continuous() ~ "continuous2"
  )
```

## Histograms