---
title: "In Class Exercise 1"
author: "ECO 6416"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Here are all the packages needed to get started.

```{r packages}
library(readxl) # to read in xlsx files
options(scipen = 999) # to remove scientific notation

sessionInfo()
```

# National Park Visitor Data

Go to Webcourses and download the `national_park_attendance.xlsx` file. Be sure to save the file in an easy to locate folder.


```{r}
parks <- read_xlsx("C:/Users/jo585802/Downloads/national_park_attendance.xlsx")
```

```{r}
head(parks)
```

```{r}
tail(parks)
```
## Dimensions of the Data

Describe the three dimensions of the data in the above table (circle correct answers)

- univariate / bivariate / multivariate

- two categorical / two quantitative / one categorical one quantitative

- cross section / time series

## Confidence Interval

```{r}
summary(parks$`Number of Visitors, 2017`)
sd(parks$`Number of Visitors, 2017`)
```

a) Give the output above, what is the best prediction of attendance at any park at the 95% level of confidence? Show the result as (low, high) and round to the nearest 100 thousand.

\newpage

b) How comfortable are you with your prediction? Justify your answer.

\vspace{1in}

## Choosing Measure of Center

Which statistic provides the best measure of center? You can use the chart below to help.

```{r}

hist(parks$`Number of Visitors, 2017`, 
     main = "Histogram of Number of Visitors, 2017",
     xlab = "Number of Visitors, 2017")


```

