---
title: "In Class Exercise 2"
author: "Joshua L Eubanks"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
```

# Wealth of 1 Percent

This dataset was obtained from FRED which shows the share of total net worth held by the top 1%. It is represented over time.

```{r wealth}

```

This gives us a range of data from the third quarter of 1989 to the first quarter of 2021

## Time Series Plot


```{r wealthplot , message=F}

```


## Time Series Regression

We can see the trend in an upwards manner, let's see what the quarterly change in slope is:

```{r wealth regression}

```

in this case, the share of wealth held by the top 1% is increasing by 0.05% each quarter.

## Prediction

Suppose we want to predict the share of wealth 2 quarters form now:

```{r}

```


# Covid cases NYT

This data was pulled for March 2020 for covid cases in the US

```{r}

```

## Plot

```{r}


```

## Regression

```{r}

```


## Prediction

What would be the predicted number of cases by April 15, 2020?

```{r}

```

## Thing 2 

On average, how many new cases of covid are there daily?

```{r}


```

