---
title: "Carnival Game"
author: "Joshua L Eubanks"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Building Example Dataset

For this exercise, I am going to create the dataset within R.

```{r data}
people <- data.frame("Height" = c(5,9,13,12,10,11,8,9,10,12,11,9,10,12,8,9,10,15,13,11) + 60,
                     "Weight" = c(140,157,205,198,162,174,150,165,170,180,170,
                                  162,165,180,160,155,165,190,185,155))
```

# Strategy 1, Guess Average Weight Every Time

```{r}
guess  <-  mean(people$Weight)
```

If we guess the average weight `r guess`, how does do we perform? If we are within the 10 pounds $\pm 5lbs$ we win \$2. Otherwise, we lose \$2.  
```{r}
people$Strat_1_Error <- people$Weight - guess

head(people, 5)
```

We can see that we created a new column which calculates how far off we were. Now let's see how much we made/loss

```{r}
people$Strat_1_GainLoss <- -2 # creating the default set of losses
people$Strat_1_GainLoss[abs(people$Strat_1_Error) <= 5] <- 2 # finding and replacing all the ones in which we won with +2 instead
```

This may seem strange, but computationally speaking, it is much faster than looping through each element.

```{r}
sum(people$Strat_1_GainLoss)
```

`r sum(people$Strat_1_GainLoss)`. Not a great day. 


# Strategy 2 Include height.

```{r}
model <- lm(Weight ~ Height, data = people)

model
```

Now that we have a relationship between height and weight, we can make more informed predictions.

```{r}
people$Strat_2_Error <- people$Weight - predict(model, newdata = data.frame(Height = people$Height))

people$Strat_2_GainLoss <- -2 # creating the default set of losses
people$Strat_2_GainLoss[abs(people$Strat_2_Error) <= 5] <- 2 # finding and replacing all the ones in which we won with +2 instead

head(people, 5)

sum(people$Strat_2_GainLoss)
```

`r sum(people$Strat_2_GainLoss)`. Much better result. What other things can we do to get higher winnings?